Elite individuals, insitutions, and economic growth accounting

### 2024-04-07 presentation/main.tex
presentation mostly done

\documentclass[10pt]{beamer}
\usepackage{pgf}
\usepackage{booktabs}
%Information to be included in the title page:
\title{Elite individuals, insitutions, and economic growth accounting}
\author{James Xu}
\institute{ECON 442, Duke University}
\date{2024}
\usefonttheme[onlymath]{serif}

@@

\begin{frame}{Summary Statistics}
    Hello
\end{frame}

@@

\begin{frame}{Limitations: data sample problems}
    Regression results do not present a clear picture of the existence of a statistical relationship between elite indicators and economic growth.
    \begin{block}{Little variation in data}
        \begin{itemize}
            \item Countries and regions with PISA test scores tend to be wealthier, more developed economies
            \item Low variance within countries over time and between countries as a result
        \end{itemize}
    \end{block}

    \begin{block}{Model specification matters}
        \begin{itemize}
            \item Including entity fixed effects makes the most significant difference versus time effects (likely due to above)
            \item Magnitudes of $math99$ is 10x larger when entity fixed effects are included; similar for $IMO$
            \item $ARWU$ and interaction terms reversed in sign compared to not including fixed effects
            \begin{itemize}
                \item Possibly due to omitted variable bias that is captured by fixed effects
            \end{itemize}
        \end{itemize}

>>>>

\documentclass[10pt]{beamer}
\usepackage{pgf}
\usepackage{booktabs}
\usepackage{array}
\usepackage{ragged2e}
% \usepackage{booktabs}
%Information to be included in the title page:
\title{Elite individuals, insitutions, and economic growth accounting}
\author{James Xu}
\institute{ECON 442, Duke University}
\date{2024}
\usefonttheme[onlymath]{serif}

@@

\begin{frame}{PISA Math Scores}
    World-wide study to evaluate 15-year-old students' performance on math, science, and reading. Testing is done usually every 3 years, starting from 2000. 2021 study was delayed to 2022 due to COVID.

    \begin{itemize}
        \item This paper uses PISA mathematics data from 2003 to 2022
        \item Countries included are mostly OECD (wealthier), with some other regions/countries participating
        \item Must use weighted means because of survey methodology $math_{c, t} = \frac{\sum_{i \in c \cap t} stu\_math_i \cdot stu\_wgt_i}{\sum_{i \in c \cap t} stu\_wgt_i}$
    \end{itemize}

    As an indicator of ``elite'' academic performance, also compute the share of students in a region/country in global 1\% of test takers (benchmark score $B_{t}$):
    \[math99_{c, t} = \frac{\sum_{i \in c \cap t} (stu\_math_i \geq B_t) \cdot stu\_wgt_i}{\sum_{i \in c \cap t} stu\_wgt_i}\]
\end{frame}

\begin{frame}{EIU Democracy Index}
    \begin{columns}
        \column{0.4\textwidth}
        Index measuring the quality of democracy around the world published by the Economist Intelligence Unit.
        \begin{itemize}
            \item Published from 2006 every 2 years until 2010, annually afterwards (use XGBoost to impute missing)
            \item 0-10 scale, where 10 is democracy and 0 is autocracy.
        \end{itemize}

        \column{0.6\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../charts/eiu_2023.jpg}
    \end{columns}

\begin{frame}{Summary Statistics}
    \begin{table}
        \caption{Summary Statistics}
        \resizebox{\linewidth}{!} {
            \begin{tabular}{lrrrrr}
            & Count & Mean & Std & Min & Max \\
            \\[-1.8ex]\hline
            \hline \\[-1.8ex]
            GDP per capita & 441.00 & 28463.99 & 25255.10 & 543.11 & 149461.79 \\
            GDP per capita growth (bps) & 440.00 & 169.06 & 452.68 & -2292.68 & 3303.05 \\
            Primary Completion & 441.00 & 90.00 & 10.25 & 51.35 & 101.95 \\
            Lower Sec. Completion & 441.00 & 77.89 & 17.14 & 29.21 & 101.97 \\
            Upper Sec. Completion & 441.00 & 62.20 & 18.48 & 18.44 & 97.40 \\
            Population & 441.00 & 35318600.19 & 59445620.16 & 34000.00 & 333287557.00 \\
            ARWU per million pop. ranked & 441.00 & 0.26 & 0.33 & 0.00 & 1.54 \\
            IMO score per log pop. & 441.00 & 4.72 & 3.49 & 0.00 & 11.79 \\
            PISA math in global 1\% & 441.00 & 0.94 & 1.46 & 0.00 & 14.64 \\
            PISA math & 441.00 & 461.93 & 56.23 & 315.96 & 574.66 \\
            \end{tabular}
        }
    \end{table}
\end{frame}

@@

\begin{frame}{Limitations: data problems}
    Regression results do not present a clear picture of the existence of a statistical relationship between elite indicators and economic growth.
    \begin{block}{Little variation in data}
        \begin{itemize}
            \item Countries and regions with PISA test scores tend to be wealthier, more developed economies
            \item Low variance within countries over time and between countries as a result
        \end{itemize}
    \end{block}

    \begin{block}{Model specification matters}
        \begin{itemize}
            \item Including entity fixed effects a significant difference versus time effects (likely due to above)
            \item Large variation in $IMO, math99$ between time effects, fixed effects, panel models
            \item Missing data (only post-2006 EIU Democracy index scores, missing school completion rates)
            \begin{itemize}
                \item XGBoost is used to fill in this data as controls are somewhat stable year-to-year, but this is not perfect
            \end{itemize}
        \end{itemize}
